---
title: "assigment_4"
author: "Enes"
date: "13 09 2020"
output:  
  html_document:
          toc: true
          toc_depth: 3
          toc_float: true
          number_sections: true
          code_folding: hide
          theme: journal 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Esoph Data

## Introduction

### Packages

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(readr)
library(ggplot2)
library(data.table)
library(dplyr)
library(ggthemes)
library(ggpubr)
library(knitr)
library(corrplot)
```


### Data Overview



```{r}
data<-data.frame()
data<-esoph

glimpse(esoph)
```

```{r}
summary(esoph)

```
## EDA for Esoph Data 

### Age Effect on Percentage of Cases

```{r,warning=FALSE,message=FALSE}
age_effect <- esoph %>% group_by(agegp) %>%
    summarise(tot_rows = n(), people_cancer = sum(ncases), people_total = sum(ncontrols),
              percentage=people_cancer*100/people_total)

ggplot(age_effect, aes(x=agegp, 
                       y=percentage, 
                       fill=agegp)) + 
                       geom_col()+
                       labs(x="Age Groups",y="Percentage of Cases", title= "Cancer Cases for each Age Groups") + 
                       theme(legend.position = "none")
```

### Alcohol Effect on Percentage of Cases

```{r,warning=FALSE,message=FALSE}
alcohol_effect <- esoph %>% group_by(alcgp) %>%
    summarise(tot_rows = n(), people_cancer = sum(ncases), people_total = sum(ncontrols),
              percentage=people_cancer*100/people_total)

ggplot(alcohol_effect, aes(x=alcgp,
                           y=percentage, 
                           fill=alcgp)) + 
        geom_col( ) +

labs(         x="Alcohol Consumption",
              y="Percentage of Cases", 
              title= "Cases for Alcohol consumption") + 
              theme(legend.position = "none")
```


### Tobocco Effect on Percentage of Cases

```{r,warning=FALSE,message=FALSE}
tobacco_effect <- esoph %>% group_by(tobgp) %>%
    summarise(tot_rows = n(), 
              people_cancer = sum(ncases), 
              people_total = sum(ncontrols),
              percentage=people_cancer*100/people_total)

ggplot(tobacco_effect, aes(x=tobgp, 
                           y=percentage, 
                           fill=tobgp)) + 
        geom_bar(stat = "identity") + 
  
  labs(x="Tobacco Consumption",
       y="Percentage of Cases", 
       title= "Cases for Tobacco consumption") + 
       theme(legend.position = "none")
```


### Comparison of Tobocco and Alcohol

```{r,warning=FALSE,message=FALSE}
tobgp_and_alcgp<- esoph %>% 
  group_by(tobgp,alcgp)%>% 
  summarise(tot_rows = n(),
              total_case = sum(ncases), 
              total_control = sum(ncontrols), 
              percentage=100*(total_case/total_control))

ggplot(tobgp_and_alcgp,aes(x=alcgp,
                       y=percentage,
                       fill=alcgp))+
  geom_col( )+
  facet_grid(tobgp ~.)+
  theme(text=element_text(size=8))+
  labs(x="Alcohol Consumption",
       y="Cancer Percentage",
       title="Alcohol Consumption and Tobocco Consumption according to Cancer Percentage") +
  theme(legend.position = "none")
```


# Young People Survey Data

## Introduction

```{r,warning=FALSE,message=FALSE}
response <- read_csv("dataset.csv")
glimpse(response)

```
## PCA for Survey Data

```{r}
fil_res <- response %>%
  select("Music":"Dancing")

fil_res=fil_res[complete.cases(fil_res),]

fil_res_pca <- princomp(as.matrix(fil_res),cor=T)
summary(fil_res_pca,loadings=TRUE)


```



```{r}
ggplot(data=data.frame(PC=1:length(fil_res_pca$sdev),
Cumulative_Variance=cumsum(fil_res_pca$sdev^2/sum(fil_res_pca$sdev^2))),
aes(x=PC,y=Cumulative_Variance)) + geom_line() +
geom_point() +
scale_y_continuous(labels = scales::percent,breaks=seq(0,1,length.out=11)) +
scale_x_continuous(breaks=seq(0,135,by=5))
```



## Multidimensional Scaling Method



```{r}
fil_res_mdsa<- response %>% select("History":"Pets")%>% drop_na()
glimpse(head(as.data.frame(fil_res_mdsa)))


```
```{r}
youth_responses_dist <- 1 - cor(fil_res_mdsa)


fil_res_mdsa <- cmdscale(youth_responses_dist,k=2)
colnames(fil_res_mdsa) <- c("x","y")
print(fil_res_mdsa)
```
```{r}
ggplot(data.frame(fil_res_mdsa),aes(x=x,y=y)) + geom_text(label=rownames(fil_res_mdsa),angle=30,size=3)+labs(title="MDS", x= "x coordinate",y="y coordinate")
```



## K-Means Method


```{r}
set.seed(58)

genre_cluster<-kmeans(fil_res_mdsa,centers=6)
mds_clusters<-
data.frame(genre=names(genre_cluster$cluster),
cluster_mds=genre_cluster$cluster) %>% arrange(cluster_mds,genre)
mds_clusters
```

```{r}
ggplot(
    data.frame(fil_res_mdsa) %>% mutate(clusters=as.factor(genre_cluster$cluster),
    genres=rownames(fil_res_mdsa)),aes(x=x,y=y)) +
    geom_text(aes(label=genres,color=clusters),angle=45,size=2) +
    geom_point(data=as.data.frame(genre_cluster$centers),aes(x=x,y=y)
)
```




## References 
1. [IE48A Lecture Notes](https://mef-bda503.github.io/archive/fall17/files/intro_to_ml.html)
2. [Data_1](https://www.kaggle.com/miroslavsabo/young-people-survey)
3. [Data_2](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/esoph)








































